<span style="color:red;">Building a house with<br />Advanced Normalization Tools (*ANTs*)</span>

### 2014-11-11 ants2015build.Rmd
ENH: new reveal style

---
title: <span style="color:red;">*ANTs*</span><br> <small>_power tools for neuroimaging_</small>
author:
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

@@

@@

<div align="center"><img width="1800" src="./figures/brian_and_nick.png" frameborder="0"></img></div>

## Long term collaborators

@@

## _ANTs_:  General purpose library for multivariate image registration, segmentation & statistical analysis tools

@@

<!--

## Built on ITK

* image registration framework

* many other classes such as:
    + `itkScalarImageToRGBImageFilter`
    + `itkLabelOverlapMeauresImageFilter`
    + `itkBSplineScatteredDataPointSetToImageFilter`
    + `itkStochasticFractalDimensionImageFilter`

## ANTs development philosophy:<br><small>Assimilate the optimal components of other approaches</small>

![](./figures/BorgFirstContact.jpg)

_"We are the Borg. Your biological and technological distinctiveness will be added to our own. Resistance is futile."_

## *The Big 5*

* <span style="color:red">image registration</span> (`antsRegistration.cxx`)

* <span style="color:red">bias correction</span> (`N4BiasFieldCorrection.cxx`)

* multivariate template construction (`antsMultivariateTemplateConstruction.sh`)

* multivariate segmentation (`Atropos.cxx`)

* cortical thickness (`KellyKaposki.cxx` and `antsCorticalThickness.sh`)

## *ANTsR*<br>for statistics and visualization

# *ANTs* optimizes mathematically well-defined <span style="color:red;">objective functions</span> guided by <span style="color:red;">prior knowledge</span>  ...

# ... including that of developers, domain experts and other colleagues ...

# plug *your ideas* into our software to gain insight into biomedical data ...

# our strong <span style="color:red;">mathematical and software engineering</span> foundation leads to near limitless opportunities for innovation in a variety of application domains
-->

# *ANTs* and the perils of circularity<br><br>[<small>"*Logical circularity in voxel-based analysis: normalization strategy may induce statistical bias*"</small>](http://www.ncbi.nlm.nih.gov/pubmed/23151955)

## *ANTs* and the perils of circularity

![](./philosophy/circularity/figures/escherANTs.png)

@@

![](./philosophy/circularity/figures/voodoo.png)

![](./philosophy/circularity/figures/fastForward.png)

![](./philosophy/circularity/figures/tbss.png)

![](./philosophy/circularity/figures/antsFlavoredTbss1.png)

![](./philosophy/circularity/figures/antsFlavoredTbss2.png)

![](./philosophy/circularity/figures/antsFlavoredTbss3.png)

![](./philosophy/circularity/figures/antsFlavoredTbss4.png)

![](./philosophy/circularity/figures/antsFlavoredTbss5.png)

![](./philosophy/circularity/figures/brianEmail.png)

![](./philosophy/circularity/figures/sleeping1.png)

![](./philosophy/circularity/figures/sleeping2.png)

![](./philosophy/circularity/figures/sleeping0.png)

![](./philosophy/circularity/figures/sleeping4.png)

![](./philosophy/circularity/figures/sleeping5.png)

![](./philosophy/circularity/figures/sleeping6.png)

![](./philosophy/circularity/figures/sleeping7.png)

![](./philosophy/circularity/figures/benefitsOfOpenScience.png)

## Command line help

<font size="5", color="lightgrey">
```
$ CreateDTICohort -h

COMMAND:
     CreateDTICohort

OPTIONS:
     -d, --image-dimensionality 2/3
     -a, --dti-atlas inputDTIAtlasFileName
     -x, --label-mask-image maskImageFileName
                            lowerThresholdFunction
     -n, --noise-sigma <noiseSigma=18>
     -p, --pathology label[<percentageChangeEig1=-0.05>,<percentageChangeAvgEig2andEig3=0.05>,<numberOfVoxels=all or percentageOfvoxels>]
     -w, --dwi-parameters [B0Image,directionFile,bvalue]
                          [B0Image,schemeFile]
     -r, --registered-population textFileWithFileNames.txt
     -o, --output [outputDirectory,fileNameSeriesRootName,<numberOfControls=10>,<numberOfExperimentals=10>]
     -h
     --help
```
</font>

## Which one is not warped?

![](./philosophy/circularity/figures/whichOneIsRegistered.png)

## _Answer_

![](./philosophy/circularity/figures/whichOneIsRegistered2.png)

![](./philosophy/circularity/figures/blobs1.png)

![](./philosophy/circularity/figures/blobs2.png)

![](./philosophy/circularity/figures/BooksteinReview1.png)
![](./philosophy/circularity/figures/BooksteinReview2.png)
![](./philosophy/circularity/figures/BooksteinReview3.png)

@@

>>>>

---
title: <span style="color:red;">Building a house with<br />Advanced Normalization Tools (*ANTs*)</span>
author: Brian B. Avants (<span style="color:blue;">PENN</span>) and <br />Nicholas J. Tustison (<span style="color:orange;">UVA</span>)
bibliography: REFERENCES.bib
logo:    logo.jpg
output:
  pdf_document:
    toc: true
    highlight: zenburn
  beamer_presentation:
    toc: true
    fig_width: 7
    fig_height: 6
  html_document:
    toc: true
    theme: readable
  revealjs_presentation:
    pandoc_args: [ "--slide-level", "2" ]
    incremental: true
    widescreen: true
    smaller: false
    theme: night
    transition: fade
    highlight: zenburn
    center: true
    self_contained: false
  ioslides_presentation:
    css: atk.css
    incremental: false
    widescreen: true
    smaller: false
    fig_width: 5
---

@@

<div align="center"><img src="./figures/brian_and_edie_sedgwick.jpg" frameborder="0"></img></div>

This talk is online at [http://stnava.github.io/ANTs2015/](http://stnava.github.io/ANTs2015/) with colored [links](http://stnava.github.io/ANTs2015/) meant to be clicked for more information.

@@

<div align="center"><img width="1800" src="./figures/bant2.png" frameborder="0"></img></div>

BBA & NT

## Long-term collaborators

@@

$+$ [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

![](./figures/lion.png)

a pride: common way of doing things

... in a competitive world ...

## Definitions

* Registration $=$ estimate an "optimal" geometric mapping between image pairs or image sets (e.g. Affine)

* <span style="color:grey;"> Similarity $=$ a function relating one image to another, given a transformation (e.g. mutual information)</span>

* Diffeomorphisms $=$ differentiable map with differentiable inverse (e.g. "silly putty", viscous fluid)

* <span style="color:grey;">Segmentation $=$ labeling tissue or anatomy in images, usually automated (e.g. K-means)</span>

* Multivariate $=$ using many voxels or measurements at once (e.g. PCA, $p >> n$ ridge regression)</span>

* <span style="color:grey;">Multiple modality $=$ using many modalities at once (e.g. DTI and T1 and BOLD)</span>

* MALF: multi-atlas label fusion - using anatomical dictionaries to label new data

* <span style="color:grey;">Solutions to challenging statistical image processing problems usually need elements from each of the above</span>

## Image mapping & perception: 1878

![](./figures/galton.png)

* Francis Galton: Can we see criminality in the face?

* (maybe he should have used *ANTs*?)

## Image mapping & biology: 1917

![](./figures/dthompson.png)

D'Arcy Thompson

## Initial scope

![](./figures/ants_initial_scope.png)

... just do a better registration (tell story) ...

## *ANTs* Lineage

![](./figures/lineage.jpg)

References: @Horn1981, @Gee1993, @Grenander1993, @Thompson2001, @Miller2002, @Shen2002, @Arnold2014, @Thirion1998, @Rueckert1999, @Fischl2012, @Ashburner2012

## Diffeomorphisms

<div align="center"><img width="1433" src="./figures/sillyputty.png" frameborder="0"></img></div>

plausible physical modeling of large, invertible deformations

"differentiable map with differentiable inverse"

## Fine-grained and flexible maps
![](./figures/highresdiffeos.jpg)

... to correct a misconception about diffeomorphisms ...

## Diffeomorphisms: image parameterization in a metric space

![](./figures/shape_appearance.png)

## General purpose library for multivariate image registration, segmentation & statistical analysis tools

@@

## *ANTs*: Beyond Registration

![](./figures/antscapabilities.jpg)

[Atropos](http://www.ncbi.nlm.nih.gov/pubmed/?term=atropos+tustison) segmentation, [N4 inhomogeneity correction](http://www.ncbi.nlm.nih.gov/pubmed/?term=N4+tustison), [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/?term=eigenanatomy+avants), [SCCAN](http://www.ncbi.nlm.nih.gov/pubmed/?term=sparse+canonical+avants), [Prior-constrained PCA](http://www.ncbi.nlm.nih.gov/pubmed/24852460), and [atlas-based label fusion](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4009425/) and [MALF](http://www.ncbi.nlm.nih.gov/pubmed/21237273) (powerful expert systems for segmentation)

## On documentation

![](figures/Chainsaw.jpg)

documentation is important

## On documentation

![](figures/blindmen.jpg)

... developers can be blind to doc deficiencies

while users are blind to what we provide!
# ANTs Impact

## ANTs Statistics

![](impact/figures/antsvisitors.png)

## ANTs & Neuroscience

We need statistical image analysis <br />at several scales in modern neuroscience

* Macro: *in vivo* structural and functional MRI

* Micro:  high-resolution post-mortem MRI links with in vivo MRI

* Nano: neuron reconstruction ...

* Solutions that are consistent across these scales have the potential
to build multi-scale feature sets or templates and provide new
insights into brain structure and function

* E.g. Parcellation constraints based on histology, tractography, function ...

* Statistical definitions of anatomy/pathology?

* Reinvention of these solutions within each lab ... can we mitigate this?

* Reduce, reuse, recycle ...

## Financial impact

* The **high error rate** in software makes testing for quality assurance critically important.

* Bugs exist **Everywhere**: so don't freak out if we admit we have bugs:  we're just being honest ( and not all devs are )

* Virtually every business in the United States now depends on software for development, production, distribution, and after-sales support of products and services.

* A 2002 NIST study estimated the direct costs to the software supply chain due to failure to identify (**successfully test for**) "bugs"

* We are better at this than most ... **not perfect**

* The estimate of direct costs compiled from industry survey data for the U.S. economy was **\$60 billion per year**

* ... this estimate did not include costs to end users such as lost business (for example, the cost of shutting down the New York Mercantile Exchange in 1998 due to a software failure).

## Reproducibility

* Consider the recent paper: [Freesurfer, laplacian, DiReCT thickness in longitudinal stroke](http://ac.els-cdn.com/S2213158214001260/1-s2.0-S2213158214001260-main.pdf?_tid=8e6f5248-335d-11e4-ba4f-00000aab0f02&acdnat=1409743944_6615d59eb5709e6bcb1aca76450f7d93)

* We "pre-wrote" a comment: [Instrumentation bias in the use and evaluation of scientific software](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3766821/?report=reader)

* With these guys:

* <img src="figures/luis.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right: 3px;"><img src="figures/torsten.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;">

## Hired by Google

<img src="figures/luis.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right: 3px;"><img src="figures/torsten.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;"><img src="figures/gang.jpg" style="height: 256px; float: middle; margin-bottom: 15px; margin-right:3px;">

# Optimal Templates

## "Optimal" templates (?)

* unbiased wrt measurement space

* space is non-linear: reference matters

* encodes prior information: still just averages (usually)

* concept extends across modalities, anatomy, temporality, etc

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.060.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.061.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.062.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.063.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.064.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.065.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.066.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.067.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.068.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.069.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.070.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.071.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.072.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.073.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.074.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.075.jpg)

## Faces, brains, whatever ...

![](templates/figures/WashUJan2011.076.jpg)

see [ANTs site](stnava.github.io/ANTs) and [ANTs Documentation page](stnava.github.io/ANTsDoc)

## Templates from ANTs Colleagues ( not us )

![](templates/figures/templates_ants_colleagues.png)

## Templates unlock modalities

![](figures/ptbp2.png)

# Evaluation

@@

<div align="center"><img width="3500" src="evaluation/figures/evalhistory.png" frameborder="0"></img></div>

@@

<!--
# New work

## RKRNS

[link](http://stnava.github.io/RKRNS/)

What mathematical encoding of language reflects both brain activity and
"our" understanding of semantics?

## RIPMARC: Imputation, strutural networks, etc.

![](figures/ripmmarc_inter.png)

MICCAI 2014, Neuroimage 2015

## Big image registration
-->

# Longitudinal processing <!--with ANTs>

<!--
## ANTs longitudinal pipeline

![](longitudinal/figures/longpipe.png)

see [link to paper](http://link.springer.com/chapter/10.1007/978-3-642-15705-9_40)

and [unbiased analysis paper](http://www.ncbi.nlm.nih.gov/pubmed/20005963)

plus [longitudinal recommendation paper](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3581852/)
-->

## `antsLongitudinalCorticalThickness.sh`

![](longitudinal/figures/longpipeex.png)

## Longitudinal TBI with lesions (segmentation)

![](longitudinal/figures/long_with_lesions_seg.png)

## Longitudinal TBI with lesions (cortical thickness)

![](longitudinal/figures/long_with_lesions_thick.png)

<!--
## ANTs longitudinal pipeline

![](longitudinal/figures/babystudy.png)
-->

# Cortical thickness with lesions

## antsCorticalThickness.sh is adaptable!

1. Register subject (or single subject template) to normal template.

2. Transform lesion mask to normal template.

3. Create additional "lesion" prior, i.e. ```SmoothImage 3 ${lesionMask} 1 ${lesionPrior} 1```.

4.  Subtract out lesion prior from all other priors and keep values $\in [0,1]$.

## Modified template spatial priors

![](lesions/figures/lesionPrior.png)

## antsCorticalThickness.sh using lesion prior

![](lesions/figures/lesionSegmentation.png)

Only change to the command call is an additional '-c WM[7]' which means "combine
the 7<sup>th</sup> prior, i.e. lesion, to the white matter for cortical thickness calculation."

# Registration & statistics:<br /> Frontiers and innovation

## multivariate statistical fields arise from fused modalities

![images at measurement fields](figures/statisticalfields.png)

*Many opportunities for statistical advancements*

## ITK+ANTs+R = <span style="color:red;">*ANTsR*</span>

## Agnostic statistics
![](./figures/antsrex.jpg)

[McMillan et al.](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3961542/)

## A Quick <span style="color:grey;">*ANTsR*</span> example
```{r loadantsr,results='hide',echo=FALSE,message=FALSE}
library(ANTsR)
```

This is an executable *ANTsR* code block - *N*-dimensional statistics to go with 
our *N*-dimensional image processing software!
```{r loadimages,message=FALSE,warning=FALSE,results='hide'}
library(ANTsR)
dim<-2
filename<-getANTsRData('r16')
img<-antsImageRead( filename , dim )
filename<-getANTsRData('r64')
img2<-antsImageRead( filename , dim )
mask<-getMask(img,50,max(img),T)
mask2<-getMask(img,150,max(img),T)
nvox<-sum( mask == 1 )
nvox2<-sum( mask2 == 1 )
```
The brain has `r nvox` voxels ...

## A Quick <span style="color:grey;">*ANTsR*</span> example

Simulate a population morphometry study - a "VBM" ...
```{r morph,results='hide'}
simnum<-10
imglist<-list()
imglist2<-list()
for ( i in 1:simnum ) {
  img1sim<-antsImageClone(img)
  img1sim[ mask==1 ]<-rnorm(nvox,mean=0.5)
  img1sim[ mask2==1 ]<-rnorm(nvox2,mean=2.0)
  img2sim<-antsImageClone(img2)
  img2sim[ mask==1 ]<-rnorm(nvox,mean=0.20)
  imglist<-lappend(imglist,img1sim)
  imglist2<-lappend(imglist2,img2sim)
}
imglist<-lappend( imglist, imglist2 )
mat<-imageListToMatrix( imglist, mask )
DX<-factor( c( rep(0,simnum), rep(1,simnum) ) )
mylmresults<-bigLMStats( lm( mat ~ DX ) )
qvals<-p.adjust( mylmresults$pval.model ) 
```
The minimum q-value is `r min(qvals)` ...

## Visualize the histograms of effects
```{r vizmorph,results='hide'}
whichvox<-qvals < 1.e-2
voxdf<-data.frame( volume=c( as.numeric( mat[,whichvox] ) ), DX=DX )
ggplot(voxdf, aes(volume, fill = DX)) + geom_density(alpha = 0.2)
```

## Visualize the anatomical distribution
```{r vizmorph2hide,results='hide',echo=FALSE}
betas<-antsImageClone(mask)
betavec<-abs(mylmresults$beta.t[1,])
betas[mask==1]<-betavec
SmoothImage(dim,betas,1.5,betas)
thresh<-paste(3.0,'x',max(betavec),sep='')
ofn<-paste(figpath,'vizmorph2.jpg',sep='')
```

```{r vizmorph2,results='hide'}
plotANTsImage(img,functional=list(betas),threshold=thresh,
  outname=ofn)
```

![](`r ofn`)

## Network visualization
see `?plotBasicNetwork`
```{r networkfn,results='hide',warning=FALSE,echo=FALSE,message=FALSE}
ofn<-paste(figpath,'network',sep='')
ofnpng<-paste(ofn,'.png',sep='')
```
```{r network,results='hide',warning=FALSE,echo=FALSE,message=FALSE,eval=F}
       mnit<-getANTsRData("mni")
       mnit<-antsImageRead(mnit,3)
       mnia<-getANTsRData("mnia")
       mnia<-antsImageRead(mnia,3)
       ThresholdImage(3,mnit,mnit,1,max(mnit))
       ImageMath(3,mnit,"FillHoles",mnit)
       cnt<-getCentroids( mnia, clustparam = 50 )
       aalcnt<-cnt$centroids[1:90,]
       brain<-renderSurfaceFunction( surfimg =list( mnit ) , alphasurf=0.1 ,smoothsval = 1.5 )
       testweights<-matrix( rep( 0, 90*90 ) ,nrow=90)
       testweights[31,37]<-1  # ant cingulate to hipp
       testweights[31,36]<-2  # ant cingulate to post cingulate
       testweights[11,65]<-3  # broca to angular 
       plotBasicNetwork( centroids = aalcnt , brain , weights=testweights )
       id<-par3d("userMatrix")
       rid<-rotate3d( id , -pi/2, 1, 0, 0 )
       rid2<-rotate3d( id , pi/2, 0, 0, 1 )
       rid3<-rotate3d( id , -pi/2, 0, 0, 1 )
       par3d(userMatrix = id ) 
       dd<-make3ViewPNG(  rid, id, rid2,  ofn )
       par3d(userMatrix = id ) 
```
![](`r ofnpng`)

## The power of *ANTs* $+$ *R* $\rightarrow$<br /> <span style="color:red;">**Reproducible imaging science**</span>

![](./figures/antsgoals.png)

... used in ["Sparse canonical correlation analysis relates network-level atrophy to multivariate cognitive measures in a neurodegenerative population"](http://www.ncbi.nlm.nih.gov/pubmed/24096125) and several upcoming ...

# Wrap-up & Conclusions

## Questions driving *ANTs* ($+$ Refs.1)

* how should we geometrically transform anatomical coordinates?

    - [syn paper](http://www.ncbi.nlm.nih.gov/pubmed/17659998) - our geometric transformation model of choice
    - [recent B-spline alternative/improvement](http://www.ncbi.nlm.nih.gov/pubmed/24409140)

* how should we measure pairwise image similarity?

    - [similarity metric evaluation](http://www.ncbi.nlm.nih.gov/pubmed/20851191) compares functions for computing rigid or affine transformations between images

* what if this pair has rgb/vector/tensor voxels?

    - [dti similarity](http://www.ncbi.nlm.nih.gov/pubmed/18041273)

* how do we extend from pairs to hundreds or thousands of pairs of images?

    - [optimal templates](http://www.ncbi.nlm.nih.gov/pubmed/?term=avants+optimal+template) and [spatial priors](http://www.ncbi.nlm.nih.gov/pubmed/24879923)
    - species specific templates/priors in [chimps](http://www.ncbi.nlm.nih.gov/pubmed/23516289) and [canines](http://www.ncbi.nlm.nih.gov/pubmed/23284904)
    - [ants "big data"](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+freesurfer)

* how do we fuse multiple modality images at the subject and population levels?

    - [ants auxiliary modality study](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+logical+circularity)

## Questions driving *ANTs* ($+$ Refs.2)

* can diffeomorphisms [improve cortical thickness measurement](http://www.ncbi.nlm.nih.gov/pubmed/?term=tustison+freesurfer)?

* how might we efficiently cluster the statistical fields that arise in image analysis?

    - [Atropos](http://www.ncbi.nlm.nih.gov/pubmed/?term=atropos+tustison) segmentation and [N4 inhomogeneity correction](http://www.ncbi.nlm.nih.gov/pubmed/?term=N4+tustison)
    - [Eigenanatomy](http://www.ncbi.nlm.nih.gov/pubmed/?term=eigenanatomy+avants) for sparse imaging-specific PCA

* how to cluster such fields when we have supervision?

    - [sparse canonical correlation analysis for neuroimaging](http://www.ncbi.nlm.nih.gov/pubmed/?term=sparse+canonical+avants)
    - [Prior-constrained PCA](http://www.ncbi.nlm.nih.gov/pubmed/24852460)
    - [atlas-based label fusion](http://www.ncbi.nlm.nih.gov/pmc/articles/PMC4009425/) and [MALF](http://www.ncbi.nlm.nih.gov/pubmed/21237273) - powerful expert systems for segmentation

* how do we implement a fully multivariate *interpretable* brain and behavior study?

    - [SCCAN for imaging & cognition](http://www.ncbi.nlm.nih.gov/pubmed/?term=sccan+avants)

* how do we extend these ideas to functional MRI & decoding?

    - [recent unpublished software](http://stnava.github.io/RKRNS/) 
    - [recent work with Ben Kandel](..not yet...)

## *ANTs* longitudinal analysis

* Longitudinal image processing issues
    - [registration induced bias](http://www.ncbi.nlm.nih.gov/pubmed/20005963)
    - [general & TBI-specific issues in longitudinal analysis](http://www.ncbi.nlm.nih.gov/pubmed/23549059)
    - [reproducibility of CBF](http://www.ncbi.nlm.nih.gov/pubmed/22517961)
    - [structure-specific analysis](http://www.ncbi.nlm.nih.gov/pubmed/22306801)

* An early study of longitudinal cortical change in ALS
    - [ALS atrophy rates](http://www.ncbi.nlm.nih.gov/pubmed/16317254)

* Extension of standard-setting ants cortical thickness pipeline to longitudinal data
    - [DynANTs (unpublished)](https://github.com/stnava/DynANTs)

## Challenges: Computational and Scientific

- Scalability: **need to fuse feature selection methods with transformation optimization**
- Scalability: **need to leverage existing ITK streaming infrastructure in application level tool**
- Domain expertise: Customizable for specific problems but sometimes not specific enough
- Rapid development: colleagues still need familiarity with compilation for latest ANTs features
- Latest theoretical advances in registration not yet wrapped for users
- Need more [Documentation](http://stnava.github.io/ANTs/)  & [testing](http://testing.psychiatry.uiowa.edu/CDash/index.php?project=ANTS) ...

## Recap

- Powerful, general-purpose, <span style="color:red;">well-evaluated</span> registration and segmentation.

- Differentiable maps with differentiable inverse <span style="color:red;">$+$ statistics in these spaces</span>

- Evaluated in multiple problem domains</span> via internal studies & open competition

- Borg philosophy: <span style="color:red;">"best of"</span> from I/O, to processing to statistical methods

- Open source, testing, many examples, consistent style, multiple platforms,  active community support ...

- Integration with *R* $+$ novel tools for prediction, decoding, high-to-low dimensional statistics.

- Collaborations with [neurodebian](http://neuro.debian.net/pkgs/ants.html), [slicer](http://www.slicer.org/), [brainsfit](https://github.com/BRAINSia/BRAINSTools), [nipype](http://nipy.sourceforge.net/nipype/), [itk](http://www.itk.org) and more ...

## Tools you can use for imaging science

- Core developers:  *B. Avants, N. Tustison, H. J. Johnson, J. T. Duda*

- Many contributors, including users ...

- Multi-platform, multi-threaded C++ [stnava.github.io/ANTs](stnava.github.io/ANTs)

- Developed in conjunction with [http://www.itk.org/](http://www.itk.org/)

- R wrapping and extension [stnava.github.io/ANTsR](stnava.github.io/ANTsR)

- rapid development, regular testing $+$ many eyes $\rightarrow$ bugs are shallow

<img style="float: right" src="figures/penn.png" />
<img style="float: left" src="figures/picsl.jpg" />

# References
